%
% Modified by Megan Patnott
% Last Change: Jan 18, 2013
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Modified version of the sample_ndthesis.tex
% by Sameer Vijay
% Last Change: Wed Jul 27 2005 14:00 CEST
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Sample Notre Dame Thesis/Dissertation
% Using Donald Peterson's ndthesis classfile
%
% Written by Jeff Squyres and Don Peterson
%
% Provided by the Information Technology Committee of
%   the Graduate Student Union
%   http://www.gsu.nd.edu/
%
% Nothing in this document is serious except the format.  :-)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is *not* a substitute for the documentation, which is included
% as a pdf file in the standard distribution, and can be obatined from
% the dtx file in the advanced distribution.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% You should *also* have a ND formatting guide to ensure that you have
% all the relevant parts, put the captions in the right place, etc.
% Just because you have this wonderful style classfile doesn't mean
% that it removes *all* the formatting onus from you.  :-)
% Although be warned that the Graduate School has been known to let
% their official formatting guide get out of date. When in doubt,
% the Microsoft Word example seemed to be the only thing kept
% consistently up-to-date in 2013, and is probably the safest thing
% to consult.
%
% You should break all of this stuff up into separate files
% (at the very least, one chapter per file) and use the \include
% command, as has been done here for chapters 1 and 2 and the appendix.
% There is also an \input command, but \include is more commonly used to
% import chapters in books and dissertations. For the differences between these
% two commands, see, e.g., 
% http://web.science.mq.edu.au/~rdale/resources/writingnotes/latexstruct.html
% or http://tex.stackexchange.com/questions/246/when-should-i-use-input-vs-include.
%
% If you compile from the command line, note that you should also have 
% a good Makefile; one that invokes LaTeX as many times as necessary 
% (up to 4) and bibtex if necessary.
%
% If you use an editor that allows you to compile from within the
% program, note that you will need to compile up to four times. Also,
% we recommend that you use pdflatex (sometimes displayed as
% LaTeX => PDF) to compile directly to pdf.
%
% If you have any suggestions, comments, questions, please send e-mail
% to: dteditor@nd.edu
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[final,numrefs,sort&compress,noinfo, twoadvisors]{nddiss2e}
% One of the options draft, review, final must be chosen.
% One of the options textrefs or numrefs should be chosen
% to specify if you want numerical or ``author-date''
% style citations.
% Other available options are:
% 10pt/11pt/12pt (available with draft only)
% twoadvisors
% noinfo (should be used when you compile the final time
%         for formal submission)
% sort (sorts multiple citations in the order that they're
%       listed in the bibliography)
% compress (compresses numerical citations, e.g. [1,2,3]
%           becomes [1-3]; has no effect when used with
%           the textrefs option)
% sort&compress (sorts and compresses numerical citations;
%           is identical to sort when used with textrefs)
\usepackage{braket}
\usepackage{multirow}

\begin{document}

\frontmatter % All the items before the first chapter go in ``frontmatter''

% Your title must be in all caps, and you must do this manually!
% Titles may be 1-4 lines long. If your title is longer than 4 lines, the class file
% may have difficulty formatting the title page.
\title{DEVELOPMENT AND APPLICATIONS OF REAL-SPACE ELECTROSTATIC INTERACTION METHODS FOR CHARGE-MULTIPOLES IN CONDENSED PHASE ENVIRONMENTS}
\author{Madan Lamichhane}
\work{Dissertation} % or \work{Thesis}
\degaward{Doctor of Philosophy} % or 
%\degaward{Master of Science \\ in \\ Subject}
\advisor{Kathie E. Newman}
\secondadvisor{J. Daniel Gezelter} % if you have two advisers are using the option twoadvisors
\department{Physics}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Front stuff
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% You must either set the copyright information or put your work in the public domain.
%\copyrightholder{Madan Lamichhane} % See template or documentation for
%\copyrightyear{2016}           % other copyright options.
%\makecopyright

% An abstract is optional for a mster's thesis, and required for a doctoral dissertation.
\begin{abstract}
Computing electrostatic interaction is the most expensive portion of a  molecular simulation. Generally, traditional and modified Ewald methods are well known for calculating electrostatic interaction in molecular dynamics (MD) communities. But these methods are computationally very expensive for larger system. Nowadays there is a growing interest in developing real space method, scales linearly with the system size. The real space method for the charge-charge interaction was originally developed by \textit{Wolf} and extended by \textit{Gezelter et al.}. In this dissertation, I present three new real-space methods: Shifted Potential (SP), Gradient-Shifted Force (GSF), and Taylor-Shifted Force (TSF), for computing electrostatic interaction between dipoles and quadrupoles. I also discuss the various static and dynamic properties evaluated using newly developed real space methods.  Additionally, I present three methods: fluctuation, perturbation, and potential of mean force (PMF) methods, for computing dielectric properties for multipolar fluids. 

The SP method is the multipolar version of the \textit{Wolf}'s electrostatic method. Similarly, the GSF and TSF method are natural extension of original Damped Shifted Force (DSF) method developed by  \textit{Fennell} and \textit{Gezelter}. The energies evaluated using SP method shows quantitative agreement with Ewald sum, therefore good choice for monte carlo (MC) simulation. The energies, forces, and torques calculated from the GSF method agrees with the Ewald's result and produce excellent conservation of energy in MD simulations. Both SP and GSF method with suitable damping parameter performs remarkably well in predicting static and dynamic properties of the liquid systems. 

For the both dipolar and quadrupolar fluids, the dielectric properties predicted using perturbation method agree with fluctuation formula. We need a correction factor to obtain the dielectric constant and susceptibility from the directly measurable polarizability from the simulation. I tabulate the correction factors for both dipolar and quadrupolar fluid for all real-space as well as Ewald method. Finally, the dielectric screening from the perturbation and fluctuation method are compared with directly measured screening factor using PMF method.  
\end{abstract}

% A dedication is optional.
%\renewcommand{\dedicationname}{DEDICATION}
\begin{dedication}
  To my late father Hari Bahadur Lamichhane and mother Dropati Lamichhane for their hard work, consistent support, love, and inspiration. 
\end{dedication}

% These are required, and must be in this order.
\tableofcontents
\listoffigures
\listoftables

% A preface is optional.
%\begin{preface}
%  I would like to preface this work with all the wonderful things that
%  Gnus have brought to our society: trees, dirt, flowers, grass,
%  lakes, and other earthly-things.  We should not forget them in our
%  daily lives.
%
%  Additionally, we should offer them food for all their hard work.  In
%  fact, Gnus work so hard that they sleep for the colder half of
%  the year.  As such, they tend to grow a little rotund.  Humans
%  should not fault them for this, as it is necessary for their
%  survival.  Indeed, many humans grow rotund on their on accord!
%\end{preface}

% It's hard to tell from the information available from the Graduate
% School in Spring 2013 whether or not an acknowledgements section is optional.
\begin{acknowledge}
First of all, I would like to express my deepest appreciations to my advisors, Professor Kathie E. Newman and Professor J. Dainel Gezelter for their persistent guidance, support, and inspiration. Without their help this dissertation would not have been possible.

I am grateful to my dissertation committee; Professor Morten Eskildsen, Professor Mark Alber, and Professor Rebecca Surman, for their valuable time, support, and suggestions. I am also thankful my research group: Dr. Joseph Michalka, Patrick Louden, Suzanne Kucera, Hemanta Bhattarai, Dr. Kelsey Stocker, and Thomas Parsons for their help, support, and valuable discussion.

I am very grateful to my wife Sumitra for her unconditional love, support, inspiration, and great patience throughout the course of my PhD research.  Without her, completing research would not this easy for me. I am also very thankful to my Parents: Hari B. Lamichhane and Dropati Lamichhane,  Brothers: Bishal and Prabin, and  Sisters: Indira and Rekha. My family has always been great source of inspiration for me. Without their support and encouragement, I would not have achieved this opportunity. I am indebted to my friends: Aboutaleb Amiri and Melinda Varga for their valuable support and discussion.

Financial support for this research was provided by the National Science Foundation (NSF) under grant CHE-1362211 and computational time was provided by the Center for Research Computing (CRC) at the University of Notre Dame.
\end{acknowledge}

\mainmatter
% Place the text body here.
%\include{chapter-one}
%Begin each chapter with \chapter{TITLE}. Chapter titles must be in all caps
%and ensuring that they are is your responsibility.

%
% Chapter 1
%

\include{chapter1}

%
% Chapter 2
%
\include{chapter2}
%
% Chapter 3

\include{chapter3}

% Chapter 4

\include{chapter4}

% Chapter 5

\include{chapter5}
% Appendix (optional)

\appendix

\include{appendixA}
%
% Back stuff
\include{appendixB}
%

% % comment out the following three lines
% if using chapter-wise bibliography

 \backmatter
 \bibliographystyle{achemso}
 %\bibliographystyle{abbrvnat} % The standard abbrvnat style should be acceptable. Also provided with both the advanced and standard
 \bibliography{dissertation}       % distributions are nddiss2e and nddiss2enoarticletitles style options.
% If you prefer to manually enter your bibliography, that is fine. Comment out the previous two lines, and enter your bibliography
% as usual. Note that if you choose this route, formatting the bibliography is your responsibility. An example is below, including the
% optional arguments necessary for author-date style citations.
%	\begin{thebibliography}{9}
%		\bibitem[Galmira(1998)]{galmira98:_gnus_milit} G.\ Galmira. Gnus and the military -- a secret conspiracy? \emph{Growing Towards Gnu}, III(7):22--183, September 1998.
%		
%		\bibitem[Ganston and Greenfield(1998)]{gnus98:_gerry_ganst} G.\ Ganston and G.\ Greenfield. \emph{Gnus and You: The Art of Being New}. volume I. Grapping Books, NY, August, 1998.
%		
%		\bibitem[Gloonston(1998)]{gloonston98:_gnuly_discov_gnus} G.\ Gloonston. Newly discovered gnus: The LoG. \emph{Growing Towards Gnu}, II(12):23---57, March 1998.
%		
%		\bibitem[Greenfield(1996)]{greenfield96:_gettin_know_gnu} G.\ Greenfield. \emph{Getting to Know Gnu}. PhD thesis, Geoffrey Garfield School of Gnus, August 1996.
%		
%		\bibitem[van Gairley(2000)]{gairley2000} G.\ van Gairley. Gnu's review. Website, 2000. \url{http://www.gairley.gnu}.
%	\end{thebibliography}

\end{document}

% End of ``dissertation.tex''
